---
title: "FIFA confederations"
author: "Dhruv Nirmal"
date: "2022-09-05"
categories: [FIFA]
image: "image.svg"
---

<h1>Motive</h1>
<div>
  <p>
  Sports occupies a vital role in our lives. It keeps us
  fit, healthy and makes us active. The secret to having a
  healthy and positive lifestyle is to have a positive mind
  and body.Apart from keeping a healthy body and mind, sports also helps
  us in alerting our senses, keeps us aware and an undaunting
  personality and outlook on life. Hence, sports has a
  multi-faceted role in our lives. Football being a major part
  of my life I decided to choose data-set related to FIFA
  audience to build my first Blog. For the purpose of analyzing and explaining 
  the data I decided to explore the data by **confederations**.  
  </p>
</div>

<h1>
Data Desription
</h1>
<div>
  <p>
  The [FIFA audience data set](https://github.com/rfordatascience/tidytuesday/blob/master/data/2018/2018-06-12) provides information about
  the different FIFA confederations (competitions) and all the
  countries that participate in them. It also shows each country's population 
  share and FIFA TV audience share in the world. I have
  analysed and to tried show a difference betwween both the
  variables for all the competitions. The dataset also provides
  vital information about each country's weighted share to GDP.
  </p>
</div>

<h1>
Variables
</h1>
<div>
  <ul>
    <li> **country**:	FIFA member country
    <li> **confederation**:	Confederation to which country belongs
    <li> **population_share**:	Country's share of global population 
    (percentage)
    <li> **tv_audience_share**:	Country's share of global world cup TV Audience 
    (percentage)
    <li> **gdp_weighted_share**:	Country's GDP-weighted audience share 
    (percentage)
  </ul>
</div>

```{r setup, include = TRUE, echo = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  eval = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 6,
  fig.pos = "H"
)
```

```{r}
library(tidyverse)
library(kableExtra)
data <- read_csv("week11_fifa_audience.csv")
```

<h2>
Data alteration
</h2>

<p>
The graph \@ref(fig:plot) shows the difference between cumulative population 
share(%) and TV audience share(%) of countries participation in different 
confederations. This is achieved by grouping the data by confederation.
</p>

```{r}
data_1 <- data %>%
  select(confederation, tv_audience_share, population_share) %>%
  pivot_longer(!confederation,
               names_to = "share_type",
               values_to = "percentage") %>%
  group_by(confederation, share_type) %>%
  summarise(contribution = sum(percentage))
```

```{r plot, fig.cap="Population share Vs TV audience share"}
ggplot(data_1,
       aes(x = reorder(confederation, contribution),
           y = contribution,
           fill = share_type)) +
  geom_bar(stat = "identity", 
           position = "dodge") +
  xlab("Confederation") +
  ylab("percentage") +
  labs(title = "Comparison between population share and tv audience share for every confederation") +
  labs(fill = "Share") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.6),
        text = element_text(size = 10))  

```

<h2>
Conclusion
</h2>
<p>
Asian countries contribute a lot to the world's population which helps us explain
as to why the participants of AFC were the majority viewers of FIFA 2010. UEFA 
being the most popular competition on club-level football, explains why their TV 
audience share is second after AFC despite having such low contribution to world's 
population. The same can be said about CONMEBOL as they are third in population 
share. OFC had almost negligible impact on both the variables.
</p>

<h2>
Variables
</h2>
<div>
  <ul>
    <li> **FIFA_members**:	Proportion of countries per confederation (percentage)
    <li> **GLOBAL POP**: Population share in the world contributed by countries per 
    confederation (percentage)
    <li> **WORLD CUP TV AUDIENCE**: TV audience share in the world contributed by 
    countries per confederation (percentage)
    <li> **GDP-WEIGHTED TV AUDIENCE**: Confederation's GDP-weighted audience 
    share (percentage)
  </ul>
</div>


<h2>
Data alteration
</h2>
<div>
<p>
For each confederation I have calculated the total contribution in population 
share, TV audience share and GDP-weighted share to show the overall contribution 
of each confederation. 
</p>
<p>
We can aggregate these numbers by each of FIFA’s six continental confederations 
and compare them to the share of FIFA members each confederation has.
</p>
</div>
```{r}
data_2 <- data %>%
  group_by(confederation) %>%
  summarise("FIFA_members" = n(),
            "GLOBAL POP" = sum(population_share),
            "WORLD CUP TV AUDIENCE" = sum(tv_audience_share),
            "GDP-WEIGHTED TV AUDIENCE" = sum(gdp_weighted_share))

table <- data_2 %>%
  summarise("FIFA_members" = (FIFA_members/sum(FIFA_members))*100,
            "GLOBAL POP" = (`GLOBAL POP`/sum(`GLOBAL POP`))*100,
            "WORLD CUP TV AUDIENCE" = (`WORLD CUP TV AUDIENCE`/sum(`WORLD CUP TV AUDIENCE`))*100,
            "GDP-WEIGHTED TV AUDIENCE" = (`GDP-WEIGHTED TV AUDIENCE`/sum(`GDP-WEIGHTED TV AUDIENCE`))*100)

table1 <- table %>%
  add_column(Confederation = c("AFC (Asia)", "CAF (Africa)", "CONCACAF (N. America)", 
                               "CONMEBOL (S. America)", "OFC (Oceania)", "UEFA (Europe)"),
             .before  = "FIFA_members") 
  
```

```{r table1, fig.cap="Each Confederation explained"}
p <- knitr::kable(table1, booktabs = TRUE, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "hold_position"), 
                full_width = F,
                font_size = 12) %>%
  kable_material_dark()
p
```

<h2>
Conclusion
</h2>

<p>
 In 2010, African countries represented about 25 percent of FIFA members but 
 (even though the World Cup was played in Africa) only about 3 percent of the 
 tournament’s GDP-weighted TV audience. By contrast, Asia and South America are 
 underrepresented in FIFA. So is UEFA if you measure it by its GDP-weighted TV 
 audience, although not if you measure it by population. You can argue, of 
 course, that it isn’t a bad thing for FIFA to give more representation to 
 developing countries. It doesn’t square with FIFA’s decision to give the 2022 
 World Cup to Qatar, which has the world’s highest per-capita GDP by some 
 estimates. (Silver, 2017)
</p>

<h1>
References:
</h1>
<div>
  <p>
    * Silver, N. (2017, April 21). How To Break FIFA. FiveThirtyEight. Retrieved September 6, 2022, from https://fivethirtyeight.com/features/how-to-break-fifa/
    * Wickham et al., (2019). Welcome to the tidyverse.
  Journal of Open Source Software, 4(43), 1686,
  https://doi.org/10.21105/joss.01686
    * Hao Zhu (2021). kableExtra: Construct Complex Table with
  'kable' and Pipe Syntax. R package version 1.3.4.
  https://CRAN.R-project.org/package=kableExtra
    * (2022). Retrieved September 6, 2022, from Wikimedia.org website: https://upload.wikimedia.org/wikipedia/commons/a/a8/World_Map_FIFA2.svg
    * Fifa Football (soccer) Confederations Emblems And Logos Editorial Stock Photo - Illustration of logotype, game: 65678403. (n.d.). Retrieved September 6, 2022, from www.dreamstime.com website: https://www.dreamstime.com/editorial-stock-photo-fifa-football-soccer-confederations-emblems-logos-collection-vector-associations-white-tablet-rusty-wooden-image65678403

    
  </p>

</div>
